% 04-application.tex

\section{Part 3 - Write a Simple Application}

\begin{frame}{Secure Timeout System Application}
    \begin{itemize}
        \item The application is a simple implementation of a Secure Timeout System.
        \item It consists of multiple tasks that simulate events, monitor activities, and handle alerts.
        \item Hardware timers are used to generate periodic interrupts for activity detection.
    \end{itemize}
\end{frame}

\begin{frame}{Task Implementation}
    \begin{itemize}
        \item \textbf{Event Task:}
        \begin{itemize}
            \item Periodically generates events that can be either user activities or suspicious activities.
            \item Uses a pseudo-random number generator to decide the type of event.
            \item Logs the generated event and updates the respective counters.
        \end{itemize}
        \item \textbf{Monitor Task:}
        \begin{itemize}
            \item Checks for user activity detection.
            \item Logs the status of user activity.
            \item Resets the user activity detection flag after logging.
        \end{itemize}
        \item \textbf{Alert Task:}
        \begin{itemize}
            \item Checks for suspicious activity detection.
            \item Logs the status of the system security.
            \item Initiates security protocols if suspicious activity is detected.
            \item Resets the suspicious activity detection flag after logging.
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{Hardware Timer Initialization}
    \begin{itemize}
        \item \textbf{Timer 0:}
        \begin{itemize}
            \item Configured to generate periodic interrupts.
            \item Interrupt handler checks for user activities and sets the user activity detection flag.
        \end{itemize}
        \item \textbf{Timer 1:}
        \begin{itemize}
            \item Configured to generate periodic interrupts.
            \item Interrupt handler checks for suspicious activities and sets the suspicious activity detection flag.
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{Implementation Details}
    \begin{itemize}
        \item \textbf{Global Variables:}
        \begin{itemize}
            \item \texttt{userActivity}, \texttt{userActivityDetection}, \texttt{suspiciousActivity}, \texttt{suspiciousActivityDetection}
        \end{itemize}
        \item \textbf{Pseudo-random Number Generator:}
        \begin{itemize}
            \item Used to simulate random events in the Event Task.
        \end{itemize}
        \item \textbf{Task Priorities:}
        \begin{itemize}
            \item Event Task has the highest priority to ensure timely event generation.
            \item Monitor Task and Alert Task have lower priorities.
        \end{itemize}
    \end{itemize}
\end{frame}
